Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-29 robdml.tex
tex

Note, in the PGD case, the sign of negative gradient of the HM objective with
respect to an adversarial perturbation $\vr$ is equivalent to the sign of
gradient for directly maximizing the hardness $\tilde{H}_\mathsf{S}$ (hence
maximizing $L_\text{T}$), \ie,
%
\begin{align}
	\Delta \vr
	=&
	\sign\big\{
		-\frac{\partial}{\partial \vr} \big\| \max(0, 
		H_\mathsf{D} - \tilde{H}_\mathsf{S} ) \big\|_2^2
	\big\}
	\\
	=&
	\sign\big\{
		2(H_\mathsf{D} - \tilde{H}_\mathsf{S})
		\frac{\partial }{\partial \vr} \tilde{H}_\mathsf{S} \big\}
	=
	\sign\big\{
		\frac{\partial}{\partial \vr} \tilde{H}_\mathsf{S}
		\big\},
\end{align}
%
assuming that $H_\mathsf{D}>\tilde{H}_\mathsf{S}$.
%
And the perturbation $\vr$ is updated as $\vr \leftarrow \vr + \alpha \Delta \vr$
by PGD, where $\alpha$ is a step size parameter.
%
Thus, the optimization of HM objective can be interpreted as directly
maximizing $\tilde{H}_\mathsf{S}$, but stops very early once it reaches
$H_\mathsf{D}$.
%
This means the model learns from an efficient adversary during adversarial
training with HM.

@@

>>>>

Note, in the PGD case, the sign of negative gradient of the HM objective with
respect to an adversarial perturbation $\vr$ is equivalent to the sign of
gradient for directly maximizing the hardness $\tilde{H}_\mathsf{S}$ (hence
maximizing $L_\text{T}$), \ie,
%
\begin{align}
	\Delta \vr
	=&
	\sign\big\{
		-\frac{\partial}{\partial \vr} \big\| \max(0, 
		H_\mathsf{D} - \tilde{H}_\mathsf{S} ) \big\|_2^2
	\big\}
	\\
	=&
	\sign\big\{
		2(H_\mathsf{D} - \tilde{H}_\mathsf{S})
		\frac{\partial }{\partial \vr} \tilde{H}_\mathsf{S} \big\}
	=
	\sign\big\{
		\frac{\partial}{\partial \vr} \tilde{H}_\mathsf{S}
		\big\},
\end{align}
%
assuming that $H_\mathsf{D}>\tilde{H}_\mathsf{S}$.
%
And the perturbation $\vr$ is updated as $\vr \leftarrow \vr + \alpha \Delta
\vr$ by PGD for $\eta$ steps with a step size $\alpha$.
%
Thus, the optimization of HM objective can be interpreted as direct
maximization of $\tilde{H}_\mathsf{S}$, which discontinues very early once it
reaches $H_\mathsf{D}$.
%
This means the model learns from an efficient adversary during adversarial
training with HM.

@@

For adversarial training of a DML model, being selective to both benign and
adversarial training samples is crucial for preventing model collapse, and
achieving good performance on both kinds of samples.
%
Hardness Manipulation is a flexible tool for us to specify such ``selection'',
and Graduate Adversary is a selection rule for adversarial examples.
%
Intra-Class Structure loss helps the model exploit more information from the
adversarial examples.
